Working Principles of Proof Assistants and Formalization of some proofs in Agda

### 2025-08-06 Mainpaper/paper.tex
presentations
minor edits

\subsection{Formal Proof language}
Proof assistants offer a dedicated programming language designed to express definitions, statements and proofs. These languages are typically Dependently typed,
and allows implicit argument . they are highly readable and rigorous. Integration with tactic scripting enables a smooth blend of automated and manual proof development [Barras et al.,
1997][Agda Team, 2023][Lean Development Group, 2024].

@@

\subsection{User Interface} 
Effective user interfaces (UIs) enable seamless interaction with the proof assistant. 
Modern UIs include integrated development environments (IDEs) or editor plugins that provide syntax highlighting, error diagnostics, real-time proof state visualization,
and automated tactic suggestion. Such interfaces significantly enhance productivity and make formal methods more accessible to a wider audience [CoqIDE, 2025][Isabelle/jEdit, 2023].
\subsection{Formal Libraries}
Proof assistants come with extensive, rigorously verified libraries that
encapsulate fundamental theories (e.g., arithmetic, set theory, algebra), supporting rapid development of
complex formalizations without reinventing foundational results.
These libraries grow continuously and provide a shared repository of reusable components vital for large-scale verification projects [Coq Standard Library, 2025][Mathlib, 2025][Isabelle Archive of Formal Proofs].

@@

\section{Upon Some Proof Assistants and Comparative Study}
\subsection{ Agda}
Agda  is a Dependently typed functional programming language and proof assistant. Extension of Martin-Lof's Type theory. (agda documentation) 
In Dependently typed programming language there is no clear distinction between types and values , and types can depend on arbitary values and may appear as results of ordinary functions. 
This feature of Dependently typed programs allows us to encode properties of values as type of proofs of that property, allowing us to use Dependent type programming as logic.  (Dependently Typed Programming in Agda)

Every agda file has a top level module which corresponds to name of the file and the main programs goes inside the top level module.
Another key component of agda is pattern matching. Pattern matching is a mechanism for analysing the structure of values.

@@

\begin{itemize}
    \item In 5.1 we did everything from scratch, starting from definitons by ourselves. For larger and more proofs doing this for every .agda code is difficult. We make use of Agda Standard Library and leverage some pre defined objects and pre proven theorems.
    \item In 5.3, For the second part of the therorem, we had take non constructive approach, which was done by using Dec from Relation.Nullary, which is the law of excluded middle.
    \item In 5.4, We tried to construct real numbers as Cauchy sequences of Rational Numbers like in \cite{BB85}, where we encountered a problem while creating a reciprocal function $\mathbb{N} \to \mathbb{Q}$. The issue being in Agda the numerator requires to be a integer, and we did not find a way to introduce the integer 0. This remains unresolved.
\end{itemize}
\newpage
% ======== References =========
\bibliographystyle{apalike} 
\bibliography{references.bib}

>>>>

\subsection{Formal Proof language}
Proof assistants offer a dedicated programming language designed to express definitions, statements and proofs. These languages are typically Dependently typed,
and allows implicit argument . they are highly readable and rigorous. Integration with tactic scripting enables a smooth blend of automated and manual proof development \cite{}

@@

\subsection{User Interface} 
Effective user interfaces (UIs) enable seamless interaction with the proof assistant. 
Modern UIs include integrated development environments (IDEs) or editor plugins that provide syntax highlighting, error diagnostics, real-time proof state visualization,
and automated tactic suggestion. Such interfaces significantly enhance productivity and make formal methods more accessible to a wider audience \cite{isabelle2023}
\subsection{Formal Libraries}
Proof assistants come with extensive, rigorously verified libraries that
encapsulate fundamental theories (e.g., arithmetic, set theory, algebra), supporting rapid development of
complex formalizations without reinventing foundational results.
These libraries grow continuously and provide a shared repository of reusable components vital for large-scale verification projects \cite{rocqstdlib:v9.0.0}

@@

\section{Upon Some Proof Assistants and Comparative Study}
\subsection{ Agda}
Agda  is a Dependently typed functional programming language and proof assistant. Extension of Martin-Lof's Type theory. (agda documentation) 
In Dependently typed programming language there is no clear distinction between types and values , and types can depend on arbitary values and may appear as results of ordinary functions. 
This feature of Dependently typed programs allows us to encode properties of values as type of proofs of that property, allowing us to use Dependent type programming as logic. \cite{norell2008agda}

Every agda file has a top level module which corresponds to name of the file and the main programs goes inside the top level module.
Another key component of agda is pattern matching. Pattern matching is a mechanism for analysing the structure of values. \cite{norell2007thesis}

@@

\begin{itemize}
    \item In 5.1 we did everything from scratch, starting from definitons by ourselves. For larger and more proofs doing this for every .agda code is difficult. We make use of Agda Standard Library and leverage some pre defined objects and pre proven theorems.
    \item In 5.3, For the second part of the therorem, we had take non constructive approach, which was done by using Dec from Relation.Nullary, which is the law of excluded middle.
    \item In 5.4, We tried to construct real numbers as Cauchy sequences of Rational Numbers like in \cite{BB85}, where we encountered a problem while creating a reciprocal function $\mathbb{N} \to \mathbb{Q}$. The issue being in Agda the numerator requires to be a integer, and we did not find a way to introduce the integer 0. This remains unresolved.
\end{itemize}
\clearpage

% ======== References =========
\bibliographystyle{apalike} 
\bibliography{references.bib}
